---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cycleTrendR

<!-- badges: start -->
[![R-CMD-check](https://github.com/PietroPiu-labstats/cycleTrendR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/PietroPiu-labstats/cycleTrendR/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

**cycleTrendR** provides adaptive trend estimation, cycle detection, Fourier harmonic
selection, bootstrap confidence intervals, change-point detection, and rolling-origin
forecasting. It supports LOESS, GAM, and GAMM models, and automatically handles
irregular sampling using the Lomb–Scargle periodogram.

The package is designed for biomedical, environmental, and engineering time-series
applications where both trend and cyclic components must be estimated robustly.

## Installation

You can install the development version of cycleTrendR like so:

```r
# install.packages("devtools")
devtools::install_github("PietroPiu-labstats/cycleTrendR")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
set.seed(1526)

  dates <- seq.Date(as.Date("2020-01-01"), by = "day", length.out = 200)
  signal <- sin(2*pi*1:200/30) + rnorm(200, 0, 0.2)

library(cycleTrendR)

res <- adaptive_cycle_trend_analysis(
  signal = signal,
  dates  = dates,
  usefourier = TRUE,
  trendmethod = "gam"
)

plot(res)

```

The output includes: Trend line, 95% bootstrap CI, Outliers, Change-points, Periodogram, Residual ACF.


What makes cycleTrendR different?


Works with irregular time series

Automatically detects cycles

Handles autocorrelation via GAMM

Provides robust uncertainty estimates

Produces publication-ready plots


When should you use it?

Biological assay drift correction

Environmental monitoring

Epidemiological seasonality

Financial cycle detection

Sensor data with autocorrelation
